---
title: "Rotavirus power calculations "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulation study design

$A\ :=$ 2-arm randomized trial.

- $N_v = 4200$ in the vaccine group $(A=1)$.
- $N_p = 3000$ in the placebo group $(A=0)$.

$S\ :=$ the immune response marker of interest measured at month 3 (observed).

- $S_1 :=$ the potential outcome of $S$ if assigned to the vaccine arm  $(A=1)$.
- $S_0 :=$ the potential outcome of $S$ if assigned to the placebo arm  $(A=0)$, not used in this analysis.


$W\ :=$ the baseline predictor

- $(W,S_1)$ ~ bivariate normal with var($W$)=var($S_1)=1$ and $\rho=$cor$(W,S_1) = 0.25, 0.5, 0.75$.

$Y\ :=$ occurrence of the rotavirus disease by 2 years post enrollment 
(i.e. after a 1.75-year follow-up period after $S$ is measured at month 3). 
 
Ideal population for inference is subjects at-risk for $Y=1$ at month 3.
(i.e., did not experience the primary rotavirus disease endpoint by month 3).  

Practically, the implementation of the method would only include subjects who are observed to attend the month 3 visit without occurrence of the rotavirus by month 3.

Marginal disease rate post marker measurement (from month 3 to month 24):

- $P(Y=1|A=0) = 0.04\times(1.75/2)$ for the placebo group. 
- $P(Y=1|A=1) = 0.01\times(1.75/2)$ for the vaccine group (75% vaccine efficacy(VE)).
 
#### Case-cohort sampling design for measuring $S$:

- $S$ measured in all $A=1$ subjects with $Y=1$ observed, 
- $S$ also measured in a simple random sample of all $A=1$ subjects. (sampling rate $=\ 25\%, 50\%, 100\%$)

For placebo recipients $A=0$ who complete follow-up with $Y=0$, cross-over $25\%, 50\%, 100\%$ to receive the vaccine and have S measured 3 months after the cross-over.  This is used only for filling in $S=S_1$, and we assume that no crossed-over subjects experience the primary rotavirus endpoint between month 24 and month 27.

#### Hypothesis of interest is 

- Null $H_0: \text{VE}_s=$VE (i.e., no modification of VE by the marker $S_1$).
- Alternative $H_1: \text{VE}_s\neq$VE (some effect modification).

Generation of $Y$: (with marginals kept)

- Null case: $Y$ generated with $W$ and $S_1$.
- Alternative case: $Y$ generated with $W$ alone.

The above set-up gives 54 scenarios:

- $\rho=(0.25,0.5,0.75)\ \times$ sampling rate $= (25\%, 50\%, 100\%)\ \times$ cross-over fraction $= (25\%, 50\%, 100\%)\ \times$ hypothesis=(null, alternative).


## Simulation results

2 examples of the 54 scenarios

```{r, echo=FALSE}
corr_S1_W= 0.25
crossover_rate_A0Y0= 0.25
crossover_rate_A1 = 0.25
al.nona <- read.csv( file=paste("~/Desktop/Peter Gilbert/2018winterRA/Results/TidyTable/corr_S1_W:",corr_S1_W,"crossover_rate_A0Y0:",crossover_rate_A0Y0,"crossover_rate_A1:",crossover_rate_A1, ".csv", sep=""))[,-1]
null.nona <- read.csv( file=paste("~/Desktop/Peter Gilbert/2018winterRA/Results/TidyTable/nullcorr_S1_W:",corr_S1_W,"crossover_rate_A0Y0:",crossover_rate_A0Y0,"crossover_rate_A1:",crossover_rate_A1, ".csv", sep=""))[,-1]
print(paste("cor(S1,W)=",corr_S1_W,", crossover rate=",crossover_rate_A0Y0,", sample rate=",crossover_rate_A1, sep=""))
print("Alternative case")
al.nona
print("Null case")
null.nona
```






